Near Optimal Behavior with Approximate Abstraction

### 2016-02-04 paper/abstract.tex
Switched from perfect to exact abstraction and other changes to introduction

The combinatorial explosion that plagues planning and \ac{RL} algorithms can be moderated using abstraction. Prohibitively difficult task representations can be condensed such that essential information is preserved, and consequently, solutions are tractably computable. However, perfect abstractions are often computationally intractable and ineffective in environments where no two situations are exactly alike. For these reasons, we investigate approximate abstractions, which treat nearly-identical situations as equivalent. In this work, we characterize sufficient conditions for approximate abstractions that preserve near optimal behavior while reducing the complexity of the original task representation. We present theoretical guarantees of the quality of behaviors derived from four classes of approximate abstractions. Additionally, we empirically demonstrate that approximate abstractions lead to reduction in task complexity and bounded loss of optimality of behavior in a variety of environments. % Environments could also be tasks/problems/etc

>>>>

The combinatorial explosion that plagues planning and \ac{RL} algorithms can be moderated using abstraction. Prohibitively difficult task representations can be condensed such that essential information is preserved, and consequently, solutions are tractably computable. However, exact abstractions, which treat fully-identical situations as equivalent, are often computationally intractable and ineffective in environments where no two situations are exactly alike. For these reasons, we investigate approximate abstractions, which treat nearly-identical situations as equivalent. In this work, we characterize sufficient conditions for approximate abstractions that preserve near optimal behavior while reducing the complexity of the original task representation. We present theoretical guarantees of the quality of behaviors derived from four classes of approximate abstractions. Additionally, we empirically demonstrate that approximate abstractions lead to reduction in task complexity and bounded loss of optimality of behavior in a variety of environments. % Environments could also be tasks/problems/etc

### 2016-02-04 paper/introduction.tex
Switched from perfect to exact abstraction and other changes to introduction

\section{Introduction}
Abstraction plays a fundamental role in learning. Through abstraction, intelligent agents may reason about only the salient features of their environment while ignoring what is irrelevant. Consequently, agents are able to solve considerably more complex problems than they would be able to without the use of abstraction. However, perfect abstractions, which treat fully-identical situations alike, require complete knowledge that is often computationally intractable to obtain. Furthermore, often no two situations are identical, so perfect abstraction is ineffective. To overcome these issues we investigate approximate abstractions that enable agents to treat sufficiently similar situations as identical. This work characterizes the impact of equating ``sufficiently similar'' states in the context of planning and \ac{RL} in \acp{MDP}. The remainder of our introduction contextualizes these intuitions in \acp{MDP}.

\dnote{Want a clause, sentence, prhase, paragraph, that transitions us from the super high level biz in the first paragraph to the second paragraph}
Abstraction in the context of planning and \ac{RL} in an \ac{MDP} takes the form of state aggregation. \dnote{rephrase and add a sentence or two that gets us to paragraph two}

@@

% Existing work on state abstraction.
Thus, a key research agenda for planning and \ac{RL} is leveraging abstraction to reduce large state spaces~\cite{andre2002state,jong2005state,dietterich2000hierarchical,Bean2011}. These methods are characterized by reducing \textit{ground} MDPs with large state spaces to \textit{abstract} MDPs with smaller state spaces by aggregating states according to some notion of equality or similarity. Existing work has characterized how aggregation of states with equal values of particular quantities fully maintains optimality~\cite{li2006towards,dean1997modelmin}. 
%However, exactly solving for these quantities is itself computationally taxing and often as difficult as solving for optimal behavior in the ground \ac{MDP}, thereby defeating the purpose of abstraction.

% Thesis: Approximate abstraction for three reasons.
However, relaxing the aggregation criteria from equality of quantities to $\epsilon$-closeness offers three benefits. First, state aggregation algorithms that satisfy these criteria employ the sort of knowledge that we expect a planning or learning algorithm to approximate without fully solving the \ac{MDP}. Second, because of their relaxed criteria, approximate abstractions can achieve greater degrees of compression than perfect abstractions. This is particularly important in environments where no two states are identical. \dnote{note this in empirical results discussion, that random is sort of like a natural environment and when eps is 0, no compression} Third, because the state aggregation criteria are relaxed to approximate equality, they are able to tune the aggressiveness of abstraction by adjusting what they consider sufficiently similar states. The thesis of this work is that aggregation of states on the basis of various approximate criteria can reduce state space size while incuring only bounded error in the resulting behavior.

>>>>

\section{Introduction}
Abstraction plays a fundamental role in learning. Through abstraction, intelligent agents may reason about only the salient features of their environment while ignoring what is irrelevant. Consequently, agents are able to solve considerably more complex problems than they would be able to without the use of abstraction. However, \textit{exact abstractions}, which treat fully-identical situations alike, require complete knowledge that is often computationally intractable to obtain. Furthermore, often no two situations are identical, so exact abstraction is ineffective. To overcome these issues we investigate \textit{approximate abstractions} that enable agents to treat sufficiently similar situations as identical. This work characterizes the impact of equating ``sufficiently similar'' states in the context of planning and \ac{RL} in \acp{MDP}. The remainder of our introduction contextualizes these intuitions in \acp{MDP}.

@@

% Existing work on state abstraction.
Thus, a key research agenda for planning and \ac{RL} is leveraging abstraction to reduce large state spaces~\cite{andre2002state,jong2005state,dietterich2000hierarchical,Bean2011}. These methods are characterized by reducing \textit{ground} \acp{MDP} with large state spaces to \textit{abstract} MDPs with smaller state spaces by aggregating states according to some notion of equality or similarity. In the context of \acp{MDP}, we understand exact abstractions as those that aggregate states with equal values of particular quantities, e.g. optimal $Q$-values. Existing work has characterized how exact abstractions can fully maintain optimality~\cite{li2006towards,dean1997modelmin}.

% Thesis: Approximate abstraction for three reasons.
However, performing approximate abstraction in \acp{MDP} by relaxing the state aggregation criteria from equality of quantities to $\epsilon$-closeness offers three benefits. First, approximate abstractions employ the sort of knowledge that we expect a planning or learning algorithm to compute without fully solving the \ac{MDP}. In contrast, exact abstractions often require solving for optimal behavior, thereby defeating the purpose of abstraction. Second, because of their relaxed criteria, approximate abstractions can achieve greater degrees of compression than exact abstractions. This is particularly important in environments where no two states are identical. \dnote{note this in empirical results discussion, that random is sort of like a natural environment and when eps is 0, no compression} Third, because the state aggregation criteria are relaxed to approximate equality, they are able to tune the aggressiveness of abstraction by adjusting what they consider sufficiently similar states. The thesis of this work is that aggregation of states on the basis of various approximate criteria can reduce state space size while incurring only bounded error in the resulting behavior.
